---
title: "Introduction to R Programming"
author: "Abby Youran QIN"
project: "Programming Program for Proleteriat"
---

# Navigation

* New file
  + R Script
  + R Notebook and R markdown ([link](https://www.google.com/url?sa=t&rct=j&q=&esrc=s&source=web&cd=&cad=rja&uact=8&ved=2ahUKEwjA-7jxycX8AhUepokEHaCNDegQFnoECAsQAQ&url=https%3A%2F%2Fwww.rstudio.com%2Fwp-content%2Fuploads%2F2015%2F02%2Frmarkdown-cheatsheet.pdf&usg=AOvVaw0EdaMWRMevvsxYfQzsnSR4) to formatting cheatsheet). 
  
* Window
  + Script Editor
  + Console & Terminal
  + Environment

* Operations
  + Run code
  + R markdown & R notebook:
    - Code chunks
    - Knit

# R Objects

Reference: [R for Cats](https://rforcats.net/)

### Numeric
```{r}
numbera <- 5 # assign value to a variable
str(numbera) # check variable type and value

numberb <- 7
numbera + numberb
```

### Character
```{r}
hi <- "hello"
str(hi)
```

### Function
```{r}
some_operation <- function(a, b){
  result <- (a+b)*(a-b)*(a/b)*(a*b)
  return(result)
}

some_operation(numbera, numberb)
```

### Vector
```{r}
numVec <- c(numbera, numberb, 3)
str(numVec)

friendlyVec <- c(hi, "hola", "nihao")
str(friendlyVec)

mixedVec <- c(numbera, 8, hi, "hallo")
str(mixedVec)
```
##### Check & change an element in the vector
```{r}
friendlyVec[2] 
friendlyVec[2] <- "Ola"
friendlyVec[2]
```
##### Subset vector
```{r}
friendlyVec1 <- friendlyVec[1:2]
friendlyVec1
```
##### Name elements in the vector
```{r}
names(friendlyVec) <- c("English", "Portuguese", "Chinese")
friendlyVec["Chinese"]
```
##### Manipulate all elements in a vector 
```{r}
# Numeric
numVec <- numVec + 1 
numVec

# Character
paste(friendlyVec, "nice to meet you!")
```


### Lists
```{r}
myList <- list(1, "a")
str(myList)
myList

mySuperList <- list("abc", 2023, myList)
str(mySuperList) # Json
```

### Data Frame
```{r}
df1 <- data.frame(friendlyVec)
df1

df2 <- data.frame(mySuperList)
df2

df3 <- data.frame(languages = c("English", "German", "Cantonese"),
                  thanks = c("thank you", "danke", "do ze"),
                  lengths = c(nchar("thank you"), nchar("danke"), nchar("do ze")))
                  # Isn't it a bit tedious? 
                  # Let's proceed to learn more techniques to work with data frames!
df3

# matrices
matrix <- matrix(c(1, 2, 3, 11, 12, 13), nrow = 2, ncol = 3)
matrix # machine learning algorithm
```

### Logical
```{r}
nay <- FALSE
yeah <- TRUE
str(nay)
str(yeah)
```

##### If Statement
```{r}
a <- 1
if (a == 1){
  print("first!")
} else if (a == 2){
  print("second!")
} else if (a == 1){
  print("First!!!")
} else if (a == 3){
  print("third!")
} else {
  print("You haven't made it to the top3")
}
```

##### While Loop
```{r}
n <- 1
while (n < 6){
  print(n)
  n = n+1
}

workshop <- "Programming Program for Proletariat"
guests <- c("Sang", "Max", "Jess", "Luhang", "Fan", "Yibing", "Abby", "Wil")

n <- 1
while (n < 7){
  print(guests[n])
  n = n+1
}

switch <- TRUE
n <- 1
while (switch){
  print(guests[n])
  n = n+1
  if (guests[n] == "Abby") {
    switch <- FALSE
  }
}
```
##### For Loop
```{r}
for (letter in workshop) {
  print(letter)
}

for (person in guests) {
  print(person)
}

for (person in guests) {
  if (person == "Abby"){
    break
  }
  print(person)
}

```
##### Compare vectors
```{r}
friendlyVec
friendlyVec1
overlap <- friendlyVec %in% friendlyVec1
overlap
friendlyVec[overlap]
friendlyVec[!overlap]
```

##### Any & All
```{r}
any(overlap == FALSE)
any(overlap == TRUE)
all(overlap == FALSE)
```

# R Libraries

##### Basic Operations
```{r eval=FALSE}
## Check your installed packages
installed.packages()

## Package installation
install.packages(c("tidyverse", "dplyr"))

## Libraries importing
library(tidyverse)
library(dplyr)
```

##### Advanced Template
```{r message=FALSE}

## Advanced template to install packages not yet installed
packages <- c("tidyverse", "dplyr") # Put all libraries you need in a vector
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages], repos = 'http://cran.us.r-project.org')
}

## Load packages
invisible(lapply(packages, library, character.only = TRUE))
```

##### Updating Packages

**Note**. You don't need to check for updates very often. Also, this is not usually a problem but in case you've been using a package extensively in your codes, you'd better check the version difference before updating that package, so that the update does not make your previous codes unusable.   
```{r eval=FALSE}
# Check packages that are not up to date
old.packages() 
# Deciding for each old package whether you'd like to update it
update.packages()

# If you only want to update a single package, just re-install it!
install.packages("dplyr")
```

##### Updating R & R Studio

R Studio: You may navigate to RStudio - Help - Check for Updates to check for available updates and update the application accordingly.  

R: You can update your R by directly downloading and installing the latest version from [CRAN](https://cran.r-project.org/). After installation, you can restart your RStudio to use the updated R. 


# R Data Frame
## Importing Data Files
```{r}
# Set working directory
setwd("/Users/abby/Documents/GitHub/ProgrammingProgramForProletariat/HelloWorld/R_1_Intro") 

# Import data
df <- read.csv("cannabis.csv")
head(df, n=3) # Check the first 3 rows (5 by default if n is not specified)
```

## Data Cleaning
##### Convert categorical variables into "factor"
```{r}
df$lang <- as.factor(df$lang)
summary(df$lang)
summary(df$source)
df$source <- as.factor(df$source)
summary(df$source)
df$reply_settings <- as.factor(df$reply_settings)
df$referenced_tweets_type <- as.factor(df$referenced_tweets_type)
```
##### Convert the date variable into "Date"
```{r}
df$created_at <- as.Date(df$created_at)
summary(df$created_at)
```
##### Subsetting a dataframe by columns
```{r}
df <- df[c(1:16)]
summary(df)
```
##### Subsetting a dataframe by rows
```{r}
dfiPhone <- subset(df, source == "Twitter for iPhone")
summary(dfiPhone)
```

## Exporting Data
```{r}
write.csv(df, "cannabis_cleaned.csv")
```

